---
title: "Initial Korean covid 19"
author: "Ashia graves"
date: "10/4/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  



  




```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The majority of the population was female, making up about 55%, while males accounted for roughly 45%. However, among those who died, nearly 65% were male and about 35% were female. This highlights a significant disparity in mortality rates, with males nearly doubling the number of deceased females.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The data reveals that the majority of the living population was female, with a ratio of 56:44. In contrast, the deceased population was predominantly male, with a ratio of 36:64. Approximately 63% of the male population had died, while only 1.4% of the female population had passed away. This indicates that the number of deceased males was more than double that of deceased females.
## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The null hypothesis suggests that the percentage of deceased males and females should be equal when compared to the living population. However, the observed data indicated that the number of deceased males was nearly double that of deceased females. The odds of females dying are 2:3, indicating that men are more likely to die than women.


## Conclusions

